Co-methylation

### 2014-09-29 Rmarkdown/comethylation.Rmd
Started introduction and background to co-methylation.

## Within-fragment co-methylation

@@

If we could sequence entire chromosomes by single reads then we could analyse co-methylation between any set of methylation loci. However, current technologies for bisulfite-sequencing mean that we are limited to studying reads that, at most, span about $200-250$ bp and so are limited to studying co-methylation of methylation loci that are very close to one another along the chromosome.

### Using pairs of methylation loci

@@

@@

>>>>

\subsection{Previous work}\label{subsec:previous work}

I define co-methylation is the dependence of DNA methylation at different methylation loci within a sample. In other words, it is the dependence or correlation structure of DNA methylation data within a sample. This is slightly different to \citeauthor{Schatz:2004cha} who defined it as, "the presence of methylation over a stretch of neighboring CpG positions" and more in line with \citeauthor{Eckhardt:2006gh} who defined it as, "the relationship between the degree of methylation over distance".

There have been several attempts to explore the relationship between methylation events at different loci within an individual. However, frequently the exact question being asked and the methodology have not been well-defined. Moreover, to the best of my knowledge there is no software available that implements any of the proposed methods.

From a biological perspective, by studying co-methylation we hope to better understand the regulation of DNA methylation. From a statistical perspective, more powerful and efficient analysis methods can be developed if this correlation structure is better understood, accounted for and exploited.

An increased understanding of co-methylation could also lead to cheaper and more efficient assays of DNA methylation. For example, we could identify methylation loci whose methylation status is highly predictive of the methylation level of a surrounding region and design an assay to interrogate those highly predictive loci. This idea is analogous to the use of tag-SNPs on whole-genome genotyping microarrays (__CITE__).

### Correlations of aggregate measurements of methylation

Most analyses of co-methylation have been based on aggregate measurements of methylation, such as $\beta$-values \citep{Eckhardt:2006gh, Cokus:2008fc, Li:2010fb, Lyko:2010dr, Hebestreit:2013ko, Lacey:2013iy, Liu:dy}. This is true of studies using both microarray data, where only these type of measurements are available, and sequencing data, where base-resolution measurements are also available. Very often the exact analysis is not well-described but it can generally be thought of as studying the autocorrelation of the $\beta$-values along the genome.

__UP TO HERE__

* Describe some analyses using autocorrelation of beta-values.
* More recently, within-fragment analyses. Describe concept (move from next section to this section).
* Other definitions/methods
  * Lister's mCG correlations
  * co-methylation across samples
  * "vertical" correlations of \citet{Capra:2014uh}

The method I developed, which is described in Section \ref{subsec:wf_cometh}, is superior because of __TODO__.

\subsection{Within-fragment co-methylation}\label{subsec:wf_cometh}

@@

If we could sequence entire chromosomes by single reads then we could analyse co-methylation between any set of methylation loci. However, current technologies for bisulfite-sequencing mean that we are limited to studying reads that, at most, span about $200-250$ bp and so are limited to studying co-methylation of methylation loci that are very close to one another along the chromosome. Obviously, we cannot study within-fragment co-methylation from microarray data since these produce measurements of methylation that are aggregated across many hundreds or thousands of cells.

\subsubsection{Using pairs of methylation loci}\label{subsec:pairs}

@@

#### Comparison of LOR to correlations

__TODO: Compare use of log-odds ratio to use of correlation coefficient and justify using LOR__

@@

## Correlations of $\beta$-values
